Sparse spectral and $p$-finite element methods for partial differential equations on spherical caps

### 2020-08-06 papers/spherical-caps/spherical-caps.tex
Further edits to paper

% Spherical Cap commands
\newcommand{\scop}{Q}
\newcommand{\scopnki}{\scop_{n,k,i}}
\newcommand{\scopmjh}{\scop_{m,j,h}}
\newcommand{\scopa}{\scop^{(a)}}
\newcommand{\scopnkia}{\scopnki^{(a)}}
\newcommand{\scopmjha}{\scopmjh^{(a)}}
\newcommand{\bigscop}{{\mathbb{Q}}}
\newcommand{\bigscopa}{\bigscop^{(a)}}
\newcommand{\bigscopN}{\bigscop_{N}}
\newcommand{\bigscopNa}{\bigscopa_{N}}
\newcommand{\bigscopNka}{\bigscopa_{N,k}}
\newcommand{\bigscopt}{\mathbb{\tilde{Q}}}
\newcommand{\bigscopta}{\bigscopt^{(a)}}
\newcommand{\bigscoptn}{\bigscopt_{n}}
\newcommand{\bigscoptna}{\bigscopta_{n}}
\newcommand{\xvec}{\mathbf{x}}
\newcommand{\ch}{Y}
\newcommand{\chki}{\ch_{k,i}}
\newcommand{\chjh}{\ch_{j,h}}
\newcommand{\chkitheta}{\ch_{k,i}(\theta)}
\newcommand{\alphaa}{\alpha^{(a)}}
\newcommand{\betaa}{\beta^{(a)}}
\newcommand{\gammaa}{\gamma^{(a)}}
\newcommand{\Wa}{W^{(a)}}
\newcommand{\bigWNa}{\mathbb{W}_N^{(a)}}
\newcommand{\ppphi}{\frac{\partial}{\partial \phi}}
\newcommand{\ppz}{\frac{\partial}{\partial z}}
\newcommand{\rhoppphi}{\rho \ppphi}
\newcommand{\pptheta}{\frac{\partial}{\partial \theta}}
\newcommand{\bigWNi}{\mathbb{W}_N^{(1)}}
\newcommand{\scopi}{\scop^{(1)}}
\newcommand{\scopnkii}{\scopnki^{(1)}}
\newcommand{\bigWNo}{\mathbb{W}_N^{(0)}}
\newcommand{\scopo}{\scop^{(0)}}
\newcommand{\scopnkio}{\scopnki^{(0)}}
\newcommand{\bigscopi}{\bigscop^{(1)}}
\newcommand{\bigscopNi}{\bigscopi_{N}}
\newcommand{\bigscopNki}{\bigscopi_{N,k}}
\newcommand{\bigscopo}{\bigscop^{(0)}}
\newcommand{\bigscopNo}{\bigscopo_{N}}
\newcommand{\bigscopNko}{\bigscopo_{N,k}}

@@

We can now define the 3D OPs for the spherical cap.
\begin{definition}\label{def:constuction}
Define the one-parameter 3D orthogonal polynomials via:
\begin{align}
	\scopnkia(x,y,z) := \genjacnmk^{(a,2k)}(x) \: \rho(x)^k \: \chki\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
\end{definition}

@@

\subsection{Jacobi matrices}

@@

\begin{lemma}\label{lemma:Qrecurrence} 
Define
\begin{align}
	\gamma_{k} :=
		\begin{cases}
			0 &\text{if } k < 0 \\
			\ch_0 &\text{if } k = 0 \\
			\half &\text{otherwise}
		\end{cases}
\end{align}
$\scopnkia(x,y,z)$ satisfy the following 3-term recurrences:
\begin{align*}
	x \: \scopnkia(x,y,z) &= \alphaa_{n,k,1} \:  \scopa_{n-1, k-1, i}(x, y, z) + \alphaa_{n,k,2} \:  \scopa_{n-1, k+1, i}(x, y, z) \nonumber \\
		& \quad \quad + \alphaa_{n,k,3} \:  \scopa_{n, k-1, i}(x, y, z) + \alphaa_{n,k,4} \:  \scopa_{n, k+1, i}(x, y, z) \nonumber \\
		& \quad \quad + \alphaa_{n,k,5} \:  \scopa_{n+1, k-1, i}(x, y, z) + \alphaa_{n,k,6} \:  \scopa_{n+1, k+1, i}(x, y, z), \\ \\
	y \: \scopnkia(x,y,z) &= \betaa_{n,k,i,1} \:  \scopa_{n-1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,2} \:  \scopa_{n-1, k+1, |i-1|}(x, y, z) \nonumber \\
		& \quad \quad + \betaa_{n,k,i,3} \:  \scopa_{n, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,4} \:  \scopa_{n, k+1, |i-1|}(x, y, z) \nonumber \\
		& \quad \quad + \betaa_{n,k,i,5} \:  \scopa_{n+1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,6} \:  \scopa_{n+1, k+1, |i-1|}(x, y, z), \\ \\
	z \: \scopnkia(x,y,z) &= \gammaa_{n,k,1} \: \scopa_{n-1, k, i}(x, y, z) + \gammaa_{n,k,2} \: \scopa_{n, k, i}(x, y, z) + \gammaa_{n,k,3} \: \scopa_{n+1, k, i}(x, y, z),
\end{align*}
for $(x,y,z) \in \Omega$, where
\begin{align*}
	\alphaa_{n,k,1} &:= \gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjacnmk^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
	\alphaa_{n,k,2} &:= \gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-2}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
	\alphaa_{n,k,3} &:=\gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+1}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
	\alphaa_{n,k,4} &:= \gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-1}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
	\alphaa_{n,k,5} &:= \gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+2}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
	\alphaa_{n,k,6} &:=\gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjacnmk^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
	\betaa_{n,k,i,j} &:= 
		\begin{cases}
			- \alphaa_{n,k,j} \quad &\text{if }(i = 0 \text{ and } j \text{ is odd}) \text{ or }(i = 1 \text{ and } j \text{ is even}) \\
			\alphaa_{n,k,j} \quad &\text{otherwise}
		\end{cases} \\	
	\gammaa_{n,k,1} &:= \beta_{n-k-1}^{(a, 2k)}, \qquad \gammaa_{n,k,2} := \alpha_{n-k}^{(a, 2k)}, \qquad \gammaa_{n,k,3} := \beta_{n-k}^{(a, 2k)}
\end{align*}
\end{lemma}

@@

Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$. Define, for $n=0,1,2,\dots$: 

@@

	\bigscoptna := 

@@

			\scopa_{n,0,0}(x,y,z) \\
			\scopa_{n,1,0}(x,y,z) \\
			\scopa_{n,1,1}(x,y,z) \\
			\scopa_{n,n,0}(x,y,z) \\
			\scopa_{n,n,1}(x,y,z)
		\end{pmatrix} \in \R^{2n+1}, 
	\bigscopta := 

@@

			\bigscopta_0 \\
			\bigscopta_1 \\
			\bigscopta_2 \\
and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to
	J_x^{(a,b,c,d)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
	J_y^{(a,b,c,d)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinition} \\
	J_z^{(a,b,c,d)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
%The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis. For example, let $a$ be a general parameter and let a function $f(x,y,z)$ defined on $\Omega$ be approximated by its expansion $f(x,y,z) = \bigscopta(x,y,z)^\top \mathbf{f}$. Then $x \: f(x,y,z)$ is approximated by $\bigscopta(x,y,z)^\top {J_x^{(a)\top}} \mathbf{f}$. In other words, ${J_x^{(a,b,c,d)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis. 
We will note here that we present these definitions of $\bigscoptna$, $\bigscopta$ and the Jacobi matrices in order to explain how we can derive the relationship for building the OPs themselves, i.e. gather a relationship \bstodoinline{is Q the right way to go or shall i derive the clenshaw matrices in Fourier mode k order?}
Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:
\begin{definition}
A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$. A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix. 
\end{definition}
For example, $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are block-tridiagonal (block-bandwidths $(1,1)$):
	J_{x/y/z}^{(a)} &= 

@@

			B^{(a)}_{x/y/z, 0} & A^{(a)}_{x/y/z, 0} & & & & \\
			C^{(a)}_{x/y/z, 1} & B^{(a)}_{x/y/z, 1} & A^{(a)}_{x/y/z, 1} & & & \\
			& C^{(a)}_{x/y/z, 2} & B^{(a)}_{x/y/z, 2} & A^{(a)}_{x/y/z, 2} & & & \\
			& & C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
For $J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
	A^{(a)}_{x,n} &:= 
			0 & A^{(a)}_{n,0,6} & 0 & & \\
			A^{(a)}_{n,1,5} & \ddots & \ddots & & \\
			& & A^{(a)}_{n,n,5} & 0 & A^{(a)}_{n,n,6} \\
	B^{(a)}_{x,n} &:= 
			0 & A^{(a)}_{n,0,4} & & \\
			A^{(a)}_{n,1,3} & \ddots & \ddots & \\
			& \ddots & \ddots & A^{(a)}_{n,n-1,4} \\
			& & A^{(a)}_{n,n,3} & 0
	C^{(a)}_{x,n} &:= 
			0 & A^{(a)}_{n,0,2} & & \\
			A^{(a)}_{n,1,1} & \ddots & \ddots & \\

@@

			& & & A^{(a)}_{n,n,1} \\

@@

	A^{(a)}_{n,k,j} &:= 

@@

	A^{(a)}_{n,0,j} &:=

@@

	A^{(a)}_{n,1,j} &:=

@@

For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:

@@

	A^{(a)}_{y,n} &:= 

>>>>

% Spherical Cap commands
\newcommand{\scop}{Q}
\newcommand{\scopnki}{\scop_{n,k,i}}
\newcommand{\scopmjh}{\scop_{m,j,h}}
\newcommand{\scopa}{\scop^{(a)}}
\newcommand{\scopnkia}{\scopnki^{(a)}}
\newcommand{\scopmjha}{\scopmjh^{(a)}}
\newcommand{\bigscop}{{\mathbb{Q}}}
\newcommand{\bigscopa}{\bigscop^{(a)}}
\newcommand{\bigscopN}{\bigscop_{N}}
\newcommand{\bigscopNa}{\bigscopa_{N}}
\newcommand{\bigscopNka}{\bigscopa_{N,k}}
\newcommand{\bigscopt}{\mathbb{\tilde{Q}}}
\newcommand{\bigscopta}{\bigscopt^{(a)}}
\newcommand{\bigscoptn}{\bigscopt_{n}}
\newcommand{\bigscoptna}{\bigscopta_{n}}
\newcommand{\xvec}{\mathbf{x}}
\newcommand{\ch}{Y}
\newcommand{\chki}{\ch_{k,i}}
\newcommand{\chjh}{\ch_{j,h}}
\newcommand{\chkitheta}{\ch_{k,i}(\theta)}
\newcommand{\alphaa}{\alpha^{(a)}}
\newcommand{\betaa}{\beta^{(a)}}
\newcommand{\gammaa}{\gamma^{(a)}}
\newcommand{\Wa}{W^{(a)}}
\newcommand{\bigWNa}{\mathbb{W}_N^{(a)}}
\newcommand{\ppphi}{\frac{\partial}{\partial \phi}}
\newcommand{\ppz}{\frac{\partial}{\partial z}}
\newcommand{\rhoppphi}{\rho \ppphi}
\newcommand{\rhozppphi}{\rho(z) \ppphi}
\newcommand{\pptheta}{\frac{\partial}{\partial \theta}}
\newcommand{\ppthetatwo}{\frac{\partial^2}{\partial \theta^2}}
\newcommand{\bigWNi}{\mathbb{W}_N^{(1)}}
\newcommand{\scopi}{\scop^{(1)}}
\newcommand{\scopnkii}{\scopnki^{(1)}}
\newcommand{\bigWNo}{\mathbb{W}_N^{(0)}}
\newcommand{\scopo}{\scop^{(0)}}
\newcommand{\scopnkio}{\scopnki^{(0)}}
\newcommand{\bigscopi}{\bigscop^{(1)}}
\newcommand{\bigscopNi}{\bigscopi_{N}}
\newcommand{\bigscopNki}{\bigscopi_{N,k}}
\newcommand{\bigscopo}{\bigscop^{(0)}}
\newcommand{\bigscopNo}{\bigscopo_{N}}
\newcommand{\bigscopNko}{\bigscopo_{N,k}}

@@

We can now define the 3D OPs for the spherical cap.
\begin{definition}\label{def:constuction}
Define the one-parameter 3D orthogonal polynomials via:
\begin{align}
	\scopnkia(x,y,z) := \genjacnmk^{(a,2k)}(z) \: \rho(z)^k \: \chki\fpr(\frac{x}{\rho(z)}, \frac{y}{\rho(z)}), \quad (x,y,z) \in \Omega, 
\end{align}
\end{definition}

@@

For clarity we state here that throughout we will regularly use coordinate transformations, mainly in any integrals over the surface $\Omega$, between the cartesian $(x,y,z)$ and the spherical coordinates $(\phi, \theta)$ -- that is, for any point $(x,y,z) \in \Omega$, let $(\phi, \theta)$ denote the angles such that:
\begin{align*}
	z &= \cos(\phi) \\
	x &= \sin(\phi) \: \cos(\theta) = \rho(z) \: \cos(\theta) \\
	y &= \sin(\phi) \: \sin(\theta) = \rho(z) \: \sin(\theta)
\end{align*}

\subsection{Jacobi matrices}\label{subsection:jacobimats}

@@

\begin{lemma}\label{lemma:Qrecurrence} 
Define
\begin{align}
	\gamma_{k} :=
		\begin{cases}
			0 &\text{if } k < 0 \\
			\ch_0 &\text{if } k = 0 \\
			\half &\text{otherwise}
		\end{cases}
\end{align}
$\scopnkia(x,y,z)$ satisfy the following recurrences:
\begin{align*}
	x \: \scopnkia(x,y,z) &= \alphaa_{n,k,1} \:  \scopa_{n-1, k-1, i}(x, y, z) + \alphaa_{n,k,2} \:  \scopa_{n-1, k+1, i}(x, y, z) \nonumber \\
		& \quad \quad + \alphaa_{n,k,3} \:  \scopa_{n, k-1, i}(x, y, z) + \alphaa_{n,k,4} \:  \scopa_{n, k+1, i}(x, y, z) \nonumber \\
		& \quad \quad + \alphaa_{n,k,5} \:  \scopa_{n+1, k-1, i}(x, y, z) + \alphaa_{n,k,6} \:  \scopa_{n+1, k+1, i}(x, y, z), \\ \\
	y \: \scopnkia(x,y,z) &= \betaa_{n,k,i,1} \:  \scopa_{n-1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,2} \:  \scopa_{n-1, k+1, |i-1|}(x, y, z) \nonumber \\
		& \quad \quad + \betaa_{n,k,i,3} \:  \scopa_{n, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,4} \:  \scopa_{n, k+1, |i-1|}(x, y, z) \nonumber \\
		& \quad \quad + \betaa_{n,k,i,5} \:  \scopa_{n+1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,6} \:  \scopa_{n+1, k+1, |i-1|}(x, y, z), \\ \\
	z \: \scopnkia(x,y,z) &= \gammaa_{n,k,1} \: \scopa_{n-1, k, i}(x, y, z) + \gammaa_{n,k,2} \: \scopa_{n, k, i}(x, y, z) + \gammaa_{n,k,3} \: \scopa_{n+1, k, i}(x, y, z),
\end{align*}
for $(x,y,z) \in \Omega$, where
\begin{align*}
	\alphaa_{n,k,1} &:= \gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjacnmk^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
	\alphaa_{n,k,2} &:= \gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-2}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
	\alphaa_{n,k,3} &:=\gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+1}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
	\alphaa_{n,k,4} &:= \gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-1}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
	\alphaa_{n,k,5} &:= \gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+2}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
	\alphaa_{n,k,6} &:=\gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjacnmk^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
	\betaa_{n,k,i,j} &:= 
		\begin{cases}
			- \alphaa_{n,k,j} \quad &\text{if }(i = 0 \text{ and } j \text{ is odd}) \text{ or }(i = 1 \text{ and } j \text{ is even}) \\
			\alphaa_{n,k,j} \quad &\text{otherwise}
		\end{cases} \\	
	\gammaa_{n,k,1} &:= \beta_{n-k-1}^{(a, 2k)}, \qquad \gammaa_{n,k,2} := \alpha_{n-k}^{(a, 2k)}, \qquad \gammaa_{n,k,3} := \beta_{n-k}^{(a, 2k)}
\end{align*}
\end{lemma}

@@

The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$. Let $N \in \N$ and define: 
\begin{align}
	\bigscopNka &:= 
		\begin{pmatrix}
			\scopa_{k,k,0}(x,y,z) \\
			\scopa_{k,k,1}(x,y,z) \\
			\vdots \\
			\scopa_{N,k,0}(x,y,z) \\
			\scopa_{N,k,1}(x,y,z)
		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
	\bigscopa_{N,0} &:= 
		\begin{pmatrix}
			\scopa_{0,0,0}(x,y,z) \\
			\vdots \\
			\scopa_{N,0,0}(x,y,z) \\
		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
	\bigscopNa &:= 
		\begin{pmatrix}
			\bigscopa_{N,0} \\
			\vdots \\
			\bigscopa_{N,N}
		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
\end{align}
and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to
\begin{align}
	J_x^{(a,b,c,d)} \: \bigscopNa(x,y,z) = x \: \bigscopNa(x,y,z), \nonumber \\
	J_y^{(a,b,c,d)} \: \bigscopNa(x,y,z) = y \: \bigscopNa(x,y,z), \label{eqn:jacobimatricesdefinition} \\
	J_z^{(a,b,c,d)} \: \bigscopNa(x,y,z) = z \: \bigscopNa(x,y,z). \nonumber
\end{align}
The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis. For example, let $a$ be a general parameter and let a function $f(x,y,z)$ defined on $\Omega$ be approximated up to order $N \in \N$ by its expansion $f(x,y,z) = \bigscopNa(x,y,z)^\top \mathbf{f}$. Then $x \: f(x,y,z)$ is approximated by $\bigscopNa(x,y,z)^\top {J_x^{(a)\top}} \mathbf{f}$. In other words, ${J_x^{(a,b,c,d)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis. Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:

\begin{definition}
A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$. A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix. 
\end{definition}

For example, $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are block-tridiagonal (block-bandwidths $(1,1)$):
\begin{align*}
	J_{x/y/z}^{(a)} &= 
		\begin{pmatrix}
			B^{(a)}_{x/y/z, 0} & A^{(a)}_{x/y/z, 0} & & & & \\
			C^{(a)}_{x/y/z, 1} & B^{(a)}_{x/y/z, 1} & A^{(a)}_{x/y/z, 1} & & & \\
			& C^{(a)}_{x/y/z, 2} & B^{(a)}_{x/y/z, 2} & A^{(a)}_{x/y/z, 2} & & & \\
			& & C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}
\end{align*}

For $J_x^{(a)}$, the sub-blocks $A^{(a)}_{x, k}, C^{(a)}_{x, k}$ have sub-block-bandwidths $(2,2)$ while the sub-blocks $B^{(a)}_{x,k}$ are zero matrices: \bstodo{fix bandwidths}
\begin{align*}
	B^{(a)}_{x,0} &= 0 \in \R^{(N+1)\times2N}, \quad B^{(a)}_{x,k} = 0 \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
	A^{(a)}_{x,k} &:= 
		\begin{pmatrix}
			A^{(a)}_{k,k,6} & & & \\
			A^{(a)}_{k+1,k,4} & \ddots & & \\
			A^{(a)}_{k+2,k,2} & \ddots & \ddots & \\
			& \ddots & \ddots & A^{(a)}_{N-1,k,6} \\
			& & A^{(a)}_{N,k,2} & A^{(a)}_{N,k,4}
		\end{pmatrix}, \quad k = 0,1,2,\dots,N \\
	A^{(a)}_{x,0} &\in \R^{(N+1)\times2N}, \quad A^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
	C^{(a)}_{x,k} &:= 
		\begin{pmatrix}
			A^{(a)}_{k,k,1} & A^{(a)}_{k,k,3} & A^{(a)}_{k,k,5} & & \\
			& \ddots & \ddots & \ddots \\
			& & \ddots & \ddots & A^{(a)}_{N-1,k,5} \\
			& & & A^{(a)}_{N,k,1} & A^{(a)}_{N,k,3}
		\end{pmatrix} \in \R^{2(N-k+1)\times2(N-k+2)}, \quad k = 1,2,\dots \\
	C^{(a)}_{x,1} &\in \R^{2N\times(N+1)}, \quad C^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k+3)} \text{ for } k = 2,3,\dots,N
\end{align*}
where for $k = 1,\dots,N, \: n = k,\dots,N$
\begin{align*}
	A^{(a)}_{n,k,j} &:= 
		\begin{pmatrix}
			\alphaa_{n,k,j} & 0 \\
			0 & \alphaa_{n,k,j}
		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
	A^{(a)}_{n,0,j} &:=
		\begin{pmatrix}
			\alphaa_{n,0,j} & 0
		\end{pmatrix} \in \R^{1\times2}, j \text{ odd} \\
	A^{(a)}_{n,1,j} &:=
		\begin{pmatrix}
			\alphaa_{n,1,j} \\
			0
		\end{pmatrix} \in \R^{2\times1}, j \text{ odd} \\
\end{align*}

For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$ while the sub-blocks $B^{(a)}_{y,k}$ are zero matrices:
\begin{align*}
	B^{(a)}_{y,0} &= 0 \in \R^{(N+1)\times2N}, \quad B^{(a)}_{y,k} = 0 \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
	A^{(a)}_{y,k} &:= 
		\begin{pmatrix}
			B^{(a)}_{k,k,6} & & & \\
			B^{(a)}_{k+1,k,4} & \ddots & & \\
			B^{(a)}_{k+2,k,2} & \ddots & \ddots & \\
			& \ddots & \ddots & B^{(a)}_{N-1,k,6} \\
			& & B^{(a)}_{N,k,2} & B^{(a)}_{N,k,4}
		\end{pmatrix}, \quad k = 0,1,2,\dots,N \\
	A^{(a)}_{y,0} &\in \R^{(N+1)\times2N}, \quad A^{(a)}_{y,k} \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
	C^{(a)}_{y,k} &:= 
		\begin{pmatrix}
			B^{(a)}_{k,k,1} & B^{(a)}_{k,k,3} & B^{(a)}_{k,k,5} & & \\
			& \ddots & \ddots & \ddots \\
			& & \ddots & \ddots & B^{(a)}_{N-1,k,5} \\
			& & & B^{(a)}_{N,k,1} & B^{(a)}_{N,k,3}
		\end{pmatrix} \in \R^{2(N-k+1)\times2(N-k+2)}, \quad k = 1,2,\dots \\
	C^{(a)}_{y,1} &\in \R^{2N\times(N+1)}, \quad C^{(a)}_{y,k} \in \R^{2(N-k+1)\times2(N-k+3)} \text{ for } k = 2,3,\dots,N
\end{align*}
where for $k = 1,\dots,N, \: n = k,\dots,N$
\begin{align*}
	B^{(a)}_{n,k,j} &:= 
		\begin{pmatrix}
			0 & \betaa_{n,k,0,j} \\
			\betaa_{n,k,1,j} & 0
		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
	B^{(a)}_{n,0,j} &:=
		\begin{pmatrix}
			0 & \betaa_{n,0,0,j}
		\end{pmatrix} \in \R^{1\times2}, j \text{ odd} \\
	B^{(a)}_{n,1,j} &:=
		\begin{pmatrix}
			0 \\
			\betaa_{n,1,1,j}
		\end{pmatrix}\in \R^{2\times1}, j \text{ odd} \\
\end{align*}

For $J_z^{(a)}$, the sub-blocks $B^{(a)}_{z,k}$ have sub-block-bandwidths $(3,3)$, while the sub-blocks $A^{(a)}_{z,k}, \: C^{(a)}_{z,k}$ are zero matrices:
\begin{align*}
	A^{(a)}_{z,0} &= 0 \in \R^{(N+1)\times2N}, \quad A^{(a)}_{z,k} = 0 \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
	C^{(a)}_{z,1} &= 0 \in \R^{2N\times(N+1)}, \quad C^{(a)}_{z,k} = 0 \in \R^{2(N-k+1)\times2(N-k+3)} \text{ for } k = 2,3,\dots,N \\
	B^{(a)}_{z,k} &:= 
		\begin{pmatrix}
			\Gamma^{(a)}_{k,k,2} & \Gamma^{(a)}_{k,k,3} & \\
			\Gamma^{(a)}_{k+1,k,1} & \ddots & \ddots & \\
			& \ddots & \ddots & \Gamma^{(a)}_{N-1,k,3} \\
			& & \Gamma^{(a)}_{N,k,1} & \Gamma^{(a)}_{N,k,2}
		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
	B^{(a)}_{z,0} &\in \R^{(N+1)\times(N+1)}, \quad B^{(a)}_{z,k} \in \R^{2(N-k+1)\times2(N-k+1)} \text{ for } k = 1,2,\dots,N
\end{align*}
where for $k = 1,\dots,N, \: n = k,\dots,N$
\begin{align*}
	\Gamma^{(a)}_{n,k,j} &:= 
		\begin{pmatrix}
			\gamma_{n,k,j} & 0 \\
			0 & \gamma_{n,k,j}
		\end{pmatrix} \in \R^{2\times2}, \\
	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}.
\end{align*}

In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated 'Clenshaw matrices'. While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the 'Clenshaw matrices' necessary for efficient function evaluation using the Clenshaw algorithm. In practice, it is simply a matter of converting coefficients between the two orderings.

Define, for $n=0,1,2,\dots$: 
\begin{align*}
	\bigscoptna := 
		\begin{pmatrix}
			\scopa_{n,0,0}(x,y,z) \\
			\scopa_{n,1,0}(x,y,z) \\
			\scopa_{n,1,1}(x,y,z) \\
			\vdots \\
			\scopa_{n,n,0}(x,y,z) \\
			\scopa_{n,n,1}(x,y,z)
		\end{pmatrix} \in \R^{2n+1}, 
	\quad \quad 
	\bigscopta := 
		\begin{pmatrix}
			\bigscopta_0 \\
			\bigscopta_1 \\
			\bigscopta_2 \\
			\vdots \\
		\end{pmatrix}
\end{align*}
and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to
\begin{align}
	\tilde J_x^{(a,b,c,d)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
	\tilde J_y^{(a,b,c,d)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
	\tilde J_z^{(a,b,c,d)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
\end{align}
where
\begin{align*}
	\tilde J_{x/y/z}^{(a)} &= 
		\begin{pmatrix}
			\tilde B^{(a)}_{x/y/z, 0} & \tilde A^{(a)}_{x/y/z, 0} & & & & \\
			\tilde C^{(a)}_{x/y/z, 1} & \tilde B^{(a)}_{x/y/z, 1} & \tilde A^{(a)}_{x/y/z, 1} & & & \\
			& \tilde C^{(a)}_{x/y/z, 2} & \tilde B^{(a)}_{x/y/z, 2} & \tilde A^{(a)}_{x/y/z, 2} & & & \\
			& & \tilde C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
\end{align*}
These matrices are then block-tridiagonal (block-bandwidths $(1,1)$).

For $\tilde J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:

@@

	\tilde A^{(a)}_{x,n} &:= 

@@

			0 & \tilde A^{(a)}_{n,0,6} & 0 & & \\
			\tilde A^{(a)}_{n,1,5} & \ddots & \ddots & & \\
			& & \tilde A^{(a)}_{n,n,5} & 0 & \tilde A^{(a)}_{n,n,6} \\
	\tilde B^{(a)}_{x,n} &:= 

@@

			0 & \tilde A^{(a)}_{n,0,4} & & \\
			\tilde A^{(a)}_{n,1,3} & \ddots & \ddots & \\
			& \ddots & \ddots & \tilde A^{(a)}_{n,n-1,4} \\
			& & \tilde A^{(a)}_{n,n,3} & 0
	\tilde C^{(a)}_{x,n} &:= 

@@

			0 & \tilde A^{(a)}_{n,0,2} & & \\
			\tilde A^{(a)}_{n,1,1} & \ddots & \ddots & \\

@@

			& & & \tilde A^{(a)}_{n,n,1} \\

@@

	\tilde A^{(a)}_{n,k,j} &:= 

@@

	\tilde A^{(a)}_{n,0,j} &:=

@@

	\tilde A^{(a)}_{n,1,j} &:=

@@

For $\tilde J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:

@@

	\tilde A^{(a)}_{y,n} &:= 
